Spin canonical divisors of log stacky curves

### 2015-06-30 stacky_write_up/spin_cring.tex
Theorem 3.5 edits

\newcommand \subhalf[1]{\frac{{#1} - 1}{2{#1}}}
\newcommand{\se}[1]{\section*{Problem #1}}
\newcommand{\halfcan}{L}
\DeclareMathOperator{\Supp}{Supp}
\DeclareMathOperator{\initial}{in}
\DeclareMathOperator{\gin}{gin}
\DeclareMathOperator{\Eff}{Eff}
\DeclareMathOperator{\sat}{sat}

@@

\todo{write below part}
For a homogeneous polynomial $f \in k[x]_{\vec{a}}$, we define the \textbf{initial term} $\initial_{\prec}(f)$ to be the monomial in the support of $f$ that is largest with respect to $\prec$, and we define $\initial_{\prec}(0)=0$.  Let $I \subseteq k[x]_{\vec{a}}$ be a homogeneous ideal.  We define the \textbf{initial ideal} $\initial_{\prec}(I)$ to be the ideal generated by $\{\initial_{\prec}(f) : f \in I\}$.  A \textbf{Gr\"obner basis} (also called a \textbf{standard basis}) for $I$ is a set $\{f_1,\dots,f_n\} \subset I$ such that 
+\[ \langle \initial_{\prec}(f_1),\dots,\initial_{\prec}(f_n) \rangle = \initial_{\prec}(I). \]

@@

%\begin{lem}
%\label{lem:minimal_quadratic}
%Suppose $k[x_1, \ldots, x_n]$ is a graded ring, not necessarily
%generated in degree 1, is equipped with a monomial ordering $\prec$.
 %Let $\phi:k[x_1, \ldots, x_n] \rightarrow B'$ be a map of graded
%rings with kernel $I'$, such that $I'$ is generated by elements $f_1,
%\ldots, f_j$, where $in_\prec(f_i) = x_i x_j,$ and no term of $f_i$
%is of the form $x_k$. Then, the elements $f_i$ determine minimal
%generators for $I$.
%\end{lem}
%\begin{proof}
%Regrade the ring $k[x_1, \ldots, x_n]$ so that all $x_i$ lie in
%degree 1. Then,
%\[
	%\phi : k[x_1, \ldots, x_n] \rightarrow B
%\]
%\noindent
%determines a map of rings, no longer necessarily graded. Consider
%the grlex monomial ordering $\prec$ on $k[x_1, \ldots, x_n]$.
%\todo{This ordering isn't quite right. 
%We need to preserve the ordering on quadratic terms}
%By
%assumption, the kernel $I$ is generated by elements whose initial
%terms have degree 2 and are all distinct. Therefore, $\dim_k (k[x_1,
%\ldots, x_n])_2 = \binom{n}{2} - j.$ This implies any set of
%generators for $I$ must have at least $j$ elements. Since $f_1,
%\ldots, f_j$ is such a set, it is minimal.
%\end{proof}
%
%\begin{rem}
%In particular, in the case of
%\[
	%\begin{tikzcd}
		%I\ar[hookrightarrow]{r}\ar[hookrightarrow]{d} & k[x_1, \ldots, x_n]
		%\ar[twoheadrightarrow]{r} \ar[hookrightarrow]{d} & B \ar[hookrightarrow]{d} \\
		%I' \ar[hookrightarrow]{r} & k[x_1, \ldots, x_n, y_1, \ldots, y_m] \ar[twoheadrightarrow]{r}& B'
	%\end{tikzcd}
%\]
%inducing
%\[
	%I/I'\hookrightarrow k[y_1, \ldots, y_m]\twoheadrightarrow B'/B
%\]
%\todo{how do we make this proof?}
%\end{rem}
%\begin{lem}
%\label{lem:minimal_quadratic}
%Suppose $k[x_1, \ldots, x_n]\hookrightarrow k[x_1, \ldots, x_n, y_1, \ldots, y_m]$ be graded ring, not necessarily
%generated in degree 1 that are equipped with monomial ordering $\prec$.
%Let $\phi: k[x_1, \ldots, x_n] \twoheadrightarrow B, \phi':k[x_1, \ldots, x_n, y_1, \ldots, y_m]\twoheadrightarrow B'$ be maps of graded
%rings with kernels $I \hookrightarrow I'$ such that the following diagram commutes:
%\[
	%\begin{tikzcd}
	%I\ar[hookrightarrow]{r}\ar[hookrightarrow]{d} & k[x_1, \ldots, x_n]\ar[twoheadrightarrow]{r} \ar[hookrightarrow]{d} & B \ar[hookrightarrow]{d}\\
	%I' \ar[hookrightarrow]{r} & k[x_1, \ldots, x_n, y_1, \ldots, y_m] \ar[twoheadrightarrow]{r}& B'
	%\end{tikzcd}
%\]
%Further suppose that $I'$ is generated over $I$ by elements $f_1,
%\ldots, f_j$ as a $k$-vector space, where $in_\prec(f_i) \in \{y_i x_j, y_i y_j\}$ and no monomial of $f_i$
%is of the form $y_i$. Then, the elements $f_i$ determine minimal
%generators for $I'$ over $I$.
%\end{lem}
%
%\begin{proof}
%Regrade the rings $k[x_1, \ldots, x_n], k[x_1, \ldots, x_n, y_1, \ldots, y_m]$ so that all $x_i$ and $y_j$ lie in
%degree 1. Then,
%\[
	%\phi : k[x_1, \ldots, x_n] \twoheadrightarrow B 
%\]
%\[
	%\phi': k[x_1, \ldots, x_n, y_1, \ldots, y_m] \twoheadrightarrow B'
%\]
%
%
%\noindent
%determines map of rings, no longer necessarily graded. 
%By assumption, $I'$ is generated over $I$ as a $k$-vector space by elements $f_1, \ldots, f_n$ whose initial terms have degree 2 and are linearly independent. By the theory of Gr\"{o}bner basis,
%\[
	%\dim_k(B'/B)_2=\dim_k(k[x_1, \ldots, x_n, y_1, \ldots, y_m]/k[x_1, \ldots, x_n])_2 - j
%\]
%so
%\[
	%\dim_k (I'/I)_2=j.
%\]
%This implies any set of
%generators for $I$ must have at least $j$ elements. Since $f_1,
%\ldots, f_j$ is such a set, it is minimal.
%\end{proof}
%
%\begin{rem}
%\label{rem:minimal_quadratic_trivial_case}
%It will be useful to use Lemma ~\ref{lem:minimal_quadratic} in the
%case $m= 0$, in which case we find explicitly that $f_1, \ldots, f_r$
%are $k$-minimal generators for $I'$ (over the trivial base vector
%space). 
%\end{rem}
%With this result, we will be able to show that the sets of generators of the ideal of relations in many of the following inductive arguments are minimal.

\ssec{Adding Points}
\label{ssec:add-points}
%In this section, we prove various inductive statements under the following setup:
%Let $\halfcan' \in \di \BQ \otimes X$ be $\BQ$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that generate $R_\halfcan$ over $\halfcan$' and initial terms of relations that generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Gr\"{o}bner Basis). \todo{Aaron: What new ideas does this description add that wasn't already described in the previous paragraph? Also, in general it's a good idea to be specific about what the most important lemmas are in the section, although I also added this to the general description of section 3.}
\todo{Pervasive change: Make sure all $\initial(I)$ are written as $\initial_\succ(I)$. We can even build this in to the macro possibly? }
Before proving three lemmas that allow us to inductively add points to the support of a divisor, we give a criterion to determine if a set of monomials generates the initial ideal of relations of $k[x_1, \ldots, x_m]\to R_D$.  This criterion will be used repeatedly to show that a given homogeneous ideal is in fact the ideal of relations.

\begin{lem}\label{lem:relations_from_generators_induction} 
Suppose $\halfcan, \halfcan' \in \di X \otimes \BQ$ with $\halfcan=\halfcan'+\frac{\alpha}{\beta}P$, such that $R_{\halfcan'}$ generated by $x_1, \ldots, x_m$
and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over 
$R_{\halfcan'}$.  Let $I'$ and $I$ be the ideals relations of $\phi':k[x_1, \ldots, x_m]\to R_{\halfcan'}$ and $\phi:k[x_1, \ldots, x_m, y_1, \ldots y_n]\to R_{\halfcan}$ respectively.
Suppose there are sets of monomials $S\subseteq R_\halfcan-R_{\halfcan'}$ and $T\subseteq R_\halfcan-(S\cup R_\halfcan')$, and a monomial ordering $\prec$ such 
\begin{enumerate}
\item $S$ forms a $k$-basis for $R_\halfcan$ over $R_\halfcan'$
\item $T\succ S\succ k[x_1, \ldots, x_n]$ {\rm(}meaning all monomials in $T$ are bigger than all monomials in $S,$ which are bigger than all monomials in $k[x_1, \ldots, x_n]${\rm)}
\item All monomials in $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ lie in 
\begin{align*}
	S \; \cup\; \langle T\rangle \; \cup \; \initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] \; \cup \; k[x_1, \ldots, x_m]
\end{align*}
\end{enumerate}
Then,
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]
	+ \langle T \rangle.
\end{align*}
\end{lem}
\begin{proof}
First, notice that $I'\subseteq I$ so 
\[
	\initial(I)\supseteq \initial(I')k[x_1, \ldots, x_m, y_1, \ldots, y_n].
\]

@@

By (3), every monomial of $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ is an element of either $S$, $\langle T\rangle$, or $\initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n].$ Therefore,
\begin{align*}
	\initial(I) & \subseteq \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] + \langle T \rangle.
\end{align*}
\end{proof}

To apply Lemma \ref{lem:relations_from_generators_induction}, we will need an appropriate monomial ordering.  \todo{Reference the lemma, instead of using previous, since things might get moved around. Aaron: Great, thanks!}
The following definition provides the necessary ordering for the Lemma \ref{lem:sat-1}.
\begin{defn}
\label{defn:graded-p-lex}
Suppose $\halfcan$ is a divisor of $X$ such that $R_\halfcan$ is generated by $x_1, \ldots x_m$.  Then we have a map $\phi: k[z_1, \ldots z_m] \rightarrow R_\halfcan, z_i\mapsto x_i$. \todo{Do we need to instantiate the $z_i$.  If so, what is the best way to do so.  Saying for indeterminates $z_1, \ldots, z_m$ seems excessive. Aaron: No, what you say is great!}  If $P$ is a point in $X$, then $\phi$ defines a graded $P$ lexicographic order {\bf{graded $P$-lex}} on $k[z_1, \ldots, z_m]$ as follows.
If $f=\prod_{i=1}^m {z_i}^{q_i}$ and $g=\prod_{i=1}^m {z_i}^{r_i}$ with $f \neq g,$ then $f\prec g$ if one of the following holds:
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord_P(f) < -\ord_P(g)$
\item $\deg(f) = \deg(g)$, $-\ord_P(f) = -\ord_P(g)$, and $q_i > r_i$ for the largest $i$ such that $q_i\ne r_i$
\end{enumerate}
\end{defn}

\todo{Perhaps note that this ordering is independent of whether we take the order relative to a line bundle}
\noindent One can easily verify graded $P$-lex is a monomial ordering in the sense defined in Cox--Little--Shea \cite[Chapter 2, $\mathsection$ 2, Definition 1]{cls:ideals-varieties-algorithms}.

We area almost ready to state Lemma ~\ref{lem:sat-1}, which will yield an inductive procedure for determining the generators and relations of $R_D$, where $D \in \di \BP^1$ is an effective divisor. While in O'Dorney \cite[Theorem 8]{dorney:canonical} gives a bound for the degree of generation of arbitrary $D \in \di \BP^1$, we obtain a much tighter bound in the case $D$ is effective. Furthermore, Lemma ~\ref{lem:sat-1} also applies to curves of genus $g > 0$. Before stating Lemma ~\ref{lem:sat-1}, we recall some notation from O'Dorney \cite{dorney:canonical}.

\todo{Peter: This sentence is very ugly.  On the other hand, I feel we need to say something about O'Dorney before the following definition, or else it is completely unmotivated.  How should we state it better?  Also note that in its current form this sentence neglects that we also extend the result to higher genus; it was too difficult to have this in though. Aaron: Do you like it any better now?}

@@

\todo{Peter: Do we want this as an [H] figure; this seems a little non-standard and can create problems with display.  Nonetheless, I copied this format to match the figure in a later section. Aaron: I'm fine with not making it an [H]}
\begin{figure}[H]\label{fig:s14/5-lattice}
\includegraphics{pics/spin-lower-approximations-pic-pics.pdf}
\caption{This figure shows each non-negative best lower approximation of $\frac{14}{5}.$ Each ``$\bullet$'' denotes a best lower approximation and each $\circ$ denotes a lattice point below $5y=14x$ which is not a best lower approximation.  Note that the non-negative best lower approximations generate the monoid of lattice points in the first quadrant satisfying  $5y \le 14x$, with the operation $(a_1, b_1)(a_2, b_2)\mapsto (a_1 + a_2, b_1 + b_2)$ .}
\end{figure}

@@

\begin{lem}
\label{lem:sat-1}
Let $X$ be a genus $g$ curve and $\halfcan' \in \di X \otimes \BQ$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}$, meaning we can choose generators $u, x_1, \ldots, x_m$ of $R_{\halfcan'}$ in degree at most $\tau$ for some $\tau\in \mathbb{N}$, with $\deg u = 1$, $\ord_P^{\halfcan'}(x_i)=0$ for all $1 \leq i \leq m$, and $\ord_P^{\halfcan'}(u) = 0$.  Suppose $\halfcan = \halfcan' + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{align}
\label{eqn:deg1-sat-ind}
	h^0(X, \lfloor{d\halfcan}\rfloor) &= h^0(X,\lfloor d\halfcan'
	\rfloor) + \left\lfloor d\frac{\alpha} {\beta} \right \rfloor &&\text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{align}
Let
\[
	0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then,

@@

\begin{align*}
	\initial_\prec(I) &= \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
										 + \langle U_i: 1 \le i \le n-1 \rangle
										 + \langle V \rangle,\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_j^{a_j}$ with $a_j\in \mathbb{Z}_{\ge 0}$ such that
\begin{enumerate}
	\item[(1)] $\sum_{j=1}^i a_j c_j \ge c_{i+1},$\\
	\item[(2)] there does not exist $(b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ with all $b_j\le a_j$ and $\sum_{j=1}^i b_jc_j\ge c_{i+1},$ \\
	\item[(3)] there does not exist $h<i$ such that $\sum_{j=1}^h y_j^{a_j}\ge c_{h+1}$.
\end{enumerate}

\item[(c)] In particular, $R_\halfcan$ is generated over $R_\halfcan'$ in degrees at most $\beta$ with $I$ generated over $I'$ in degrees at most $\max(2\beta, \beta + \tau)$ (where $\tau = \max(1, \max_{1\le i, m}(\deg(x_i)))$).  \todo{Peter: Part (c) is immediate from (a) and (b), so maybe it shouldn't be its own section.  On the other hand, the other parts already have a lot of information written down (especial part (b)), so its not clear where it should be put.  Maybe in a remark after the lemma? Aaron: I like including it in the lemma. I think it's really the most important point of this, so it's fine to include it here. It's also ok to just include this as a statement after parts a and b, if we don't particularly care about referencing part c.}
%\item If $\gin_\prec(I')=\initial_\prec(I')$ then $\gin_\prec(I)=\initial_\prec(I)$.
%\item[(c)] A Gr\"{o}bner basis with the initial terms $y_i x_j$, $y_i y_j$,
%	and $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
%	\frac{\alpha}{\beta} \rfloor} + 1}$ (really the subset of these with positive subscripts) as shown above minimally
%generates $I'$.
\end{enumerate}
\end{lem}
\begin{proof}
{\bf Part (a):}
By
Equation ~\ref{eqn:deg1-sat-ind}, for any $d
\in \mathbb{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

%together with $\initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$.

@@

			%in_\prec(I) &= in_\prec(I')k[x_2, \ldots, x_m, y_1, y_2] \\

@@

%\prec(I)$. Thus, it is sufficient to show that $J\subseteq \initial_\prec(I)$. For $j\ge 2$, we 

@@

%\noindent
%Furthermore, $R'$ satisfies the hypothesis of Lemma ~\ref
%{lem:raise-stacky-order}(d), which asserts that for all $i \in J$, $e_i >
%\deg z$ for any generator $z$ of $R'$ and if $R' = k[y, x]/I'$
%(as defined in Lemma ~\ref{lem:raise-stacky-order}(c)) then
%$in_\prec(I')$ is minimally generated by products of two monomials.

>>>>

\newcommand \subhalf[1]{\frac{{#1} - 1}{2{#1}}}
\newcommand{\se}[1]{\section*{Problem #1}}
\newcommand{\halfcan}{L}
\DeclareMathOperator{\Supp}{Supp}
\DeclareMathOperator{\initial}{in_\prec}
\DeclareMathOperator{\gin}{gin}
\DeclareMathOperator{\Eff}{Eff}
\DeclareMathOperator{\sat}{sat}

@@

\todo{write below part}
For a homogeneous polynomial $f \in k[x]_{\vec{a}}$, we define the \textbf{initial term} $\initial(f)$ to be the monomial in the support of $f$ that is largest with respect to $\prec$, and we define $\initial(0)=0$.  Let $I \subseteq k[x]_{\vec{a}}$ be a homogeneous ideal.  We define the \textbf{initial ideal} $\initial(I)$ to be the ideal generated by $\{\initial(f) : f \in I\}$.  A \textbf{Gr\"obner basis} (also called a \textbf{standard basis}) for $I$ is a set $\{f_1,\dots,f_n\} \subset I$ such that 
+\[ \langle \initial(f_1),\dots,\initial(f_n) \rangle = \initial(I). \]

@@

%\begin{lem}
%\label{lem:minimal_quadratic}
%Suppose $k[x_1, \ldots, x_n]$ is a graded ring, not necessarily
%generated in degree 1, is equipped with a monomial ordering $\prec$.
 %Let $\phi:k[x_1, \ldots, x_n] \rightarrow B'$ be a map of graded
%rings with kernel $I'$, such that $I'$ is generated by elements $f_1,
%\ldots, f_j$, where $\initial(f_i) = x_i x_j,$ and no term of $f_i$
%is of the form $x_k$. Then, the elements $f_i$ determine minimal
%generators for $I$.
%\end{lem}
%\begin{proof}
%Regrade the ring $k[x_1, \ldots, x_n]$ so that all $x_i$ lie in
%degree 1. Then,
%\[
	%\phi : k[x_1, \ldots, x_n] \rightarrow B
%\]
%\noindent
%determines a map of rings, no longer necessarily graded. Consider
%the grlex monomial ordering $\prec$ on $k[x_1, \ldots, x_n]$.
%\todo{This ordering isn't quite right. 
%We need to preserve the ordering on quadratic terms}
%By
%assumption, the kernel $I$ is generated by elements whose initial
%terms have degree 2 and are all distinct. Therefore, $\dim_k (k[x_1,
%\ldots, x_n])_2 = \binom{n}{2} - j.$ This implies any set of
%generators for $I$ must have at least $j$ elements. Since $f_1,
%\ldots, f_j$ is such a set, it is minimal.
%\end{proof}
%
%\begin{rem}
%In particular, in the case of
%\[
	%\begin{tikzcd}
		%I\ar[hookrightarrow]{r}\ar[hookrightarrow]{d} & k[x_1, \ldots, x_n]
		%\ar[twoheadrightarrow]{r} \ar[hookrightarrow]{d} & B \ar[hookrightarrow]{d} \\
		%I' \ar[hookrightarrow]{r} & k[x_1, \ldots, x_n, y_1, \ldots, y_m] \ar[twoheadrightarrow]{r}& B'
	%\end{tikzcd}
%\]
%inducing
%\[
	%I/I'\hookrightarrow k[y_1, \ldots, y_m]\twoheadrightarrow B'/B
%\]
%\todo{how do we make this proof?}
%\end{rem}
%\begin{lem}
%\label{lem:minimal_quadratic}
%Suppose $k[x_1, \ldots, x_n]\hookrightarrow k[x_1, \ldots, x_n, y_1, \ldots, y_m]$ be graded ring, not necessarily
%generated in degree 1 that are equipped with monomial ordering $\prec$.
%Let $\phi: k[x_1, \ldots, x_n] \twoheadrightarrow B, \phi':k[x_1, \ldots, x_n, y_1, \ldots, y_m]\twoheadrightarrow B'$ be maps of graded
%rings with kernels $I \hookrightarrow I'$ such that the following diagram commutes:
%\[
	%\begin{tikzcd}
	%I\ar[hookrightarrow]{r}\ar[hookrightarrow]{d} & k[x_1, \ldots, x_n]\ar[twoheadrightarrow]{r} \ar[hookrightarrow]{d} & B \ar[hookrightarrow]{d}\\
	%I' \ar[hookrightarrow]{r} & k[x_1, \ldots, x_n, y_1, \ldots, y_m] \ar[twoheadrightarrow]{r}& B'
	%\end{tikzcd}
%\]
%Further suppose that $I'$ is generated over $I$ by elements $f_1,
%\ldots, f_j$ as a $k$-vector space, where $\initial(f_i) \in \{y_i x_j, y_i y_j\}$ and no monomial of $f_i$
%is of the form $y_i$. Then, the elements $f_i$ determine minimal
%generators for $I'$ over $I$.
%\end{lem}
%
%\begin{proof}
%Regrade the rings $k[x_1, \ldots, x_n], k[x_1, \ldots, x_n, y_1, \ldots, y_m]$ so that all $x_i$ and $y_j$ lie in
%degree 1. Then,
%\[
	%\phi : k[x_1, \ldots, x_n] \twoheadrightarrow B 
%\]
%\[
	%\phi': k[x_1, \ldots, x_n, y_1, \ldots, y_m] \twoheadrightarrow B'
%\]
%
%
%\noindent
%determines map of rings, no longer necessarily graded. 
%By assumption, $I'$ is generated over $I$ as a $k$-vector space by elements $f_1, \ldots, f_n$ whose initial terms have degree 2 and are linearly independent. By the theory of Gr\"{o}bner basis,
%\[
	%\dim_k(B'/B)_2=\dim_k(k[x_1, \ldots, x_n, y_1, \ldots, y_m]/k[x_1, \ldots, x_n])_2 - j
%\]
%so
%\[
	%\dim_k (I'/I)_2=j.
%\]
%This implies any set of
%generators for $I$ must have at least $j$ elements. Since $f_1,
%\ldots, f_j$ is such a set, it is minimal.
%\end{proof}
%
%\begin{rem}
%\label{rem:minimal_quadratic_trivial_case}
%It will be useful to use Lemma ~\ref{lem:minimal_quadratic} in the
%case $m= 0$, in which case we find explicitly that $f_1, \ldots, f_r$
%are $k$-minimal generators for $I'$ (over the trivial base vector
%space). 
%\end{rem}
%With this result, we will be able to show that the sets of generators of the ideal of relations in many of the following inductive arguments are minimal.

\ssec{Adding Points}
\label{ssec:add-points}
%In this section, we prove various inductive statements under the following setup:
%Let $\halfcan' \in \di \BQ \otimes X$ be $\BQ$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that generate $R_\halfcan$ over $\halfcan$' and initial terms of relations that generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Gr\"{o}bner Basis). \todo{Aaron: What new ideas does this description add that wasn't already described in the previous paragraph? Also, in general it's a good idea to be specific about what the most important lemmas are in the section, although I also added this to the general description of section 3.}
\todo{Pervasive change: Make sure all $\initial(I)$ are written as $\initial(I)$. We can even build this in to the macro possibly? }
Before proving three lemmas that allow us to inductively add points to the support of a divisor, we give a criterion to determine if a set of monomials generates the initial ideal of relations of $k[x_1, \ldots, x_m]\to R_D$.  This criterion will be used repeatedly to show that a given homogeneous ideal is in fact the ideal of relations.

\begin{lem}\label{lem:relations_from_generators_induction} 
Suppose $\halfcan, \halfcan' \in \di X \otimes \BQ$ with $\halfcan=\halfcan'+\frac{\alpha}{\beta}P$, such that $R_{\halfcan'}$ generated by $x_1, \ldots, x_m$
and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over 
$R_{\halfcan'}$.  Let $I'$ and $I$ be the ideals relations of $\phi':k[x_1, \ldots, x_m]\to R_{\halfcan'}$ and $\phi:k[x_1, \ldots, x_m, y_1, \ldots y_n]\to R_{\halfcan}$ respectively.
Suppose there are sets of monomials $S\subseteq R_\halfcan-R_{\halfcan'}$ and $T\subseteq R_\halfcan-(S\cup R_\halfcan')$, and a monomial ordering $\prec$ such 
\begin{enumerate}
\item $S$ forms a $k$-basis for $R_\halfcan$ over $R_\halfcan'$
\item $T\succ S\succ k[x_1, \ldots, x_n]$ {\rm(}meaning all monomials in $T$ are bigger than all monomials in $S,$ which are bigger than all monomials in $k[x_1, \ldots, x_n]${\rm)}
\item All monomials in $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ lie in 
\begin{align*}
	S \; \cup\; \langle T\rangle \; \cup \; \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] \; \cup \; k[x_1, \ldots, x_m]
\end{align*}
\end{enumerate}
Then,
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]
	+ \langle T \rangle.
\end{align*}
\end{lem}
\begin{proof}
First, notice that $I'\subseteq I$ so 
\[
	\initial(I)\supseteq \initial(I')k[x_1, \ldots, x_m, y_1, \ldots, y_n].
\]

@@

By (3), every monomial of $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ is an element of either $S$, $\langle T\rangle$, or $\initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n].$ Therefore,
\begin{align*}
	\initial(I) & \subseteq \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] + \langle T \rangle.
\end{align*}
\end{proof}

To apply Lemma \ref{lem:relations_from_generators_induction}, we will need an appropriate monomial ordering.
The following definition provides the necessary ordering for the Lemma \ref{lem:sat-1}.
\begin{defn}
\label{defn:graded-p-lex}
Suppose $\halfcan$ is a divisor of $X$ such that $R_\halfcan$ is generated by $x_1, \ldots x_m$.  Then we have a map $\phi: k[z_1, \ldots z_m] \rightarrow R_\halfcan, z_i\mapsto x_i$. \todo{Do we need to instantiate the $z_i$.  If so, what is the best way to do so.  Saying for indeterminates $z_1, \ldots, z_m$ seems excessive. Aaron: No, what you say is great!}  If $P$ is a point in $X$, then $\phi$ defines a graded $P$ lexicographic order {\bf{graded $P$-lex}} on $k[z_1, \ldots, z_m]$ as follows.
If $f=\prod_{i=1}^m {z_i}^{q_i}$ and $g=\prod_{i=1}^m {z_i}^{r_i}$ with $f \neq g,$ then $f\prec g$ if one of the following holds:
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord_P(f) < -\ord_P(g)$
\item $\deg(f) = \deg(g)$, $-\ord_P(f) = -\ord_P(g)$, and $q_i > r_i$ for the largest $i$ such that $q_i\ne r_i$
\end{enumerate}
\end{defn}

\begin{rem}\label{rem:graded-P-lex-independent-of-line-bundle}
Observe that Definition \ref{defn:graded-p-lex} remains the same if we replace $-\ord_P$ with $-\ord_P^{L'}$ for any divisor $L'$ of $X$.
\end{rem}

\noindent One can easily verify graded $P$-lex is a monomial ordering in the sense defined in Cox--Little--Shea \cite[Chapter 2, $\mathsection$ 2, Definition 1]{cls:ideals-varieties-algorithms}.

We area almost ready to state Lemma ~\ref{lem:sat-1}, which will yield an inductive procedure for determining the generators and relations of $R_D$, where $D \in \di \BP^1$ is an effective $
\mathbb{Q}$-divisor. Whereas O'Dorney considers arbitrary $\mathbb{Q}$ divisors in $Div(\mathbb{P}^1)$ \cite[Theorem 8]{dorney:canonical}, we restrict attention to effective divisors and in Lemma ~\ref{lem:sat-1} we are able to obtain much tighter bounds.  Moreover, Lemma ~\ref{lem:sat-1} also extends to curves of genus $g > 0$. Before stating Lemma ~\ref{lem:sat-1}, we recall some notation from O'Dorney \cite{dorney:canonical}.

@@

\todo{Peter: Do we want this as an [H] figure; this seems a little non-standard and can create problems with display.  Nonetheless, I copied this format to match the figure in a later section. Aaron: I'm fine with not making it an [H]}
\begin{figure}\label{fig:s14/5-lattice}
\includegraphics{pics/spin-lower-approximations-pic-pics.pdf}
\caption{This figure shows each non-negative best lower approximation of $\frac{14}{5}.$ Each ``$\bullet$'' denotes a best lower approximation and each $\circ$ denotes a lattice point below $5y=14x$ which is not a best lower approximation.  Note that the non-negative best lower approximations generate the monoid of lattice points in the first quadrant satisfying  $5y \le 14x$, with the operation $(a_1, b_1)(a_2, b_2)\mapsto (a_1 + a_2, b_1 + b_2)$ .}
\end{figure}

@@

\begin{lem}
\label{lem:sat-1}
Let $X$ be a genus $g$ curve and $\halfcan' \in \di X \otimes \BQ$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}$, meaning we can choose generators $u, x_1, \ldots, x_m$ of $R_{\halfcan'}$ in degree at most $\tau$ for some $\tau\in \mathbb{N}$, with $\deg u = 1$, $\ord_P^{\halfcan'}(x_i)=0$ for all $1 \leq i \leq m$, and $\ord_P^{\halfcan'}(u) = 0$.  Suppose $\halfcan = \halfcan' + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{align}
\label{eqn:deg1-sat-ind}
	h^0(X, \lfloor{d\halfcan}\rfloor) &= h^0(X,\lfloor d\halfcan'
	\rfloor) + \left\lfloor d\frac{\alpha} {\beta} \right \rfloor &&\text{ for all } d \in \mathbb{
	N}.
\end{align}
Let
\[
	0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then,

@@

\begin{align*}
	\initial(I) &= \initial(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
										 + \langle U_i: 1 \le i \le n-1 \rangle
										 + \langle V \rangle,\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_j^{a_j}$ with $a_j\in \mathbb{Z}_{\ge 0}$ such that
\begin{enumerate}
	\item[(*)] $\sum_{j=1}^i a_j c_j \ge c_{i+1},$\\
	\item[(**)] there does not exist $(b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ with all $b_j\le a_j$ and $\sum_{j=1}^i b_jc_j\ge c_{i+1},$ \\
	\item[(***)] there does not exist $r<i$ such that $\sum_{j=1}^r y_j^{a_j}\ge c_{r+1}$.
\end{enumerate}

\item[(c)] In particular, $R_\halfcan$ is generated over $R_\halfcan'$ in degrees at most $\beta$ with $I$ generated over $I'$ in degrees at most $\max(2\beta, \beta + \tau)$ (where $\tau = \max(1, \max_{1\le i, m}(\deg(x_i)))$).
\end{enumerate}
\end{lem}
\begin{proof}
{\bf Part (a):}
By
Equation ~\ref{eqn:deg1-sat-ind}, for any $d
\in \mathbb{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

%together with $\initial(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$.

@@

			%\initial(I) &= \initial(I')k[x_2, \ldots, x_m, y_1, y_2] \\

@@

%\prec(I)$. Thus, it is sufficient to show that $J\subseteq \initial(I)$. For $j\ge 2$, we 

@@

%\noindent
%Furthermore, $R'$ satisfies the hypothesis of Lemma ~\ref
%{lem:raise-stacky-order}(d), which asserts that for all $i \in J$, $e_i >
%\deg z$ for any generator $z$ of $R'$ and if $R' = k[y, x]/I'$
%(as defined in Lemma ~\ref{lem:raise-stacky-order}(c)) then
%$\initial(I')$ is minimally generated by products of two monomials.
